---
title: "Exploration using FinancialModelingR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploration using FinancialModelingR package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Initialization

`FinancialModelingR` package contains several data sets corresponding to commodities, especially soybeans. These data sets are publicly available and are refreshed in different frequencies. The following data sets will be used as examples to show the usage of `FinancialModelingR` for exploring the effects of data sets of different granularities on the daily commodity/stock price:

- Soybean crop progress
- Soybean weekly exports
- Soybean WASDE reports

Functions used for reading and preprocessing these data sets has also been provided in `FinancialModelingR` package. These functions are not discussed in this notebook. In this notebook the preprocessed data sets are directly loaded and used for exploration.

## Crop Progress

```{r warning=F, message=F}
library(FinancialModelingR)
soybeanCropProgress2019 <- readRDS("preprocessed_data/soybeanCropProgress2019.Rds")
```

## Loading Private Data

Private data sets used:

- Soybean July 2020 contract prices
- Tweets related to China, soybeans and farmers from \@realdonaldtrump Twitter handle

### Price Data

```{r fig.width = 7, fig.height = 4}
contractsForJuly2020 <- readRDS("preprocessed_data/contractsForJuly2020.Rds")
print(head(contractsForJuly2020))
```

### Tweets

This module merges the DTM with price to prepare for plotting.

```{r}
# DTM = document term matrix
tweet_dtm_df <- readRDS("preprocessed_data/tweet_df.Rds")
print(head(tweet_dtm_df))
price_tweet_dtm_df <- merge(tweet_dtm_df,
                            contractsForJuly2020[, c("Date", "july_2020_Close")],
                            by = "Date")
```

# Plotting

## Daily Data

### Soybean Tweets Example

```{r fig.width = 7, fig.height = 4}
create_corr_plot(independent_var_names = NULL,
                 dependent_var_names = "july_2020_Close",
                 df = price_tweet_dtm_df,
                 remove_cols = c("amp", "get", "want", "many", "will", "just",
                                 "want", "can", "realdonaldtrump", "created_at",
                                 "made", "going", "must", "make", "Date"))

price_tweet_dtm_df$china_jobs <- apply(price_tweet_dtm_df[, -1], 1, function(row) {
  as.integer((row["china"] * row["jobs"]) != 0)
})
create_corr_plot(independent_var_names = c("china", "trade", "money",
                                           "deal", "tariffs",
                                           "economy", "currency", "china_jobs"),
                 dependent_var_names = "july_2020_Close",
                 df = price_tweet_dtm_df)
```

## Weekly Data

### Soybean Crop Progress Example

```{r fig.width = 7, fig.height = 4, warning = F, message=F}
plot_price_vs_weekly_series(df1_progress = soybeanCropProgress2019,
                            df2_contracts = contractsForJuly2020) +
  labs(title = "Crop_progress_19 vs March Contract price")
```

### Soybean Exports Example

```{r fig.width = 7, fig.height = 4}
df_top_export <- readRDS("preprocessed_data/top_10_export_countries.Rds")
df_top_export <- merge(contractsForJuly2020, df_top_export, by = "Date")
ind_col <- c("Weekly_Exports", "CMY_Outstanding_Sales", "CMY_Gross_New_Sales",
             "CMY_Net_Sales" ,"CMY_Total_Commitment")
corr_mat_weekly <- create_corr_plot(
  independent_var_names = ind_col,
  dependent_var_names = colnames(df_top_export)[
    grep("Close",colnames(df_top_export))],
  df = df_top_export,
  selected_countries = unique(df_top_export$Country),
  country_var = "Country",
  remove_countries = c("GRAND TOTAL", "KNOWN"))
```

## Monthly Data

### WASDE Example

```{r fig.width=7, fig.height=4}
soybeanWASDE_clean <- readRDS("preprocessed_data/soybeanWASDE_clean.Rds")
plot_monthly_data(soybeanWASDE_clean)
```
