---
title: "Preprocessing using FinancialModelingR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preprocessing data using FinancialModelingR package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Introduction

`FinancialModelingR` package contains several data sets corresponding to commodities, especially soybeans. The following preprocessed data sets are already included:

- Soybean crop progress
- Soybean weekly exports
- Soybean WASDE reports

This notebook discusses how `FinancialModelingR` can be used to preprocess raw input files for the following types of data sets:

- Soybean crop progress
- Soybean weekly exports
- Soybean WASDE reports
- Commodity price
- Twitter API output

# Commodity Futures Price

```{r fig.width=7, fig.height=4, warning=F, message=F}
library(FinancialModelingR)
library(png)
library(grid)
img <- readPNG("private_data/july2020.PNG")
grid.raster(img)
```

This module can be used to read contracts price from a given xlsx file that contains daily open, close, low and high prices of contracts. This data is private and has not been included in the package. The structure of the file is as shown above. The function `read_price` can be used to read the file and create columns with change in price. It can also be used to subset the price data by start date and to rename the price columns by appending a given prefix.

```{r}
contractsForJuly2020 <- FinancialModelingR::read_price(
  in_file = "private_data/ActiveSoybeanContractsforJuly2020.xlsx", delta_price = T,
  add_delta = T, subset = T, subset_min_date = "2017-01-01",
  rename_price_columns = T, rename_prefix = "july_2020_", skip_lines = 3)
contractsForJuly2020$Date <- as.Date(contractsForJuly2020$Date, "%Y-%m-%d")
saveRDS(contractsForJuly2020, "preprocessed_data/contractsForJuly2020.Rds")
```

# Tweets *Rework*

```{r fig.width=7, fig.height=4, warning=F, message=F}
img <- readPNG("private_data/realdonaldtrump.PNG")
grid.raster(img)
```

This section uses document term matrix (DTM) created from Excel file containing tweets. The DTM is aggregated in daily level.

```{r}
tweet_df <- readRDS("private_data/text_features.Rds")
colnames(tweet_df)[1] <- "Date"
sds <- sapply(tweet_df[, 2:ncol(tweet_df)], sd)
remove_cols <- 1 + which(sds == 0)
if(length(remove_cols) > 0) {
  tweet_df <- tweet_df[, -remove_cols]
}
saveRDS(tweet_df, "preprocessed_data/tweet_df.Rds")
```

# Export Sales Report *Rework*

```{r}
library(dplyr)
data("soybeanExports", package = "FinancialModelingR")
competitors <- c("ARGENTINA", "BRAZIL")
df_total_export <- soybeanExports %>% group_by(Country) %>%
  summarize(Total_Export = sum(Weekly_Exports, na.rm = T))
top_countries <- head(x = df_total_export$Country[
  order(df_total_export$Total_Export, decreasing = TRUE)], n = 10)
selected_countries <- c(competitors, top_countries)
df_top_export <- soybeanExports[sapply(
  soybeanExports$Country, function(country) country %in% selected_countries), ]
saveRDS(df_top_export, "preprocessed_data/top_10_export_countries.Rds")
```

# Crop Progress *Rework*

```{r}
data("soybeanCropProgressUSA2019", package = "FinancialModelingR")
soybeanCropProgress2019$WEEK.ENDING <-
  as.Date(soybeanCropProgress2019$WEEK.ENDING, "%Y-%m-%d")
saveRDS(soybeanCropProgress2019, "preprocessed_data/soybeanCropProgress2019.Rds")
```

# WASDE Report *Rework*

```{r}
data("soybeanCombinedWASDE")
soybeanWASDE_clean <- clean_wasde(combined_data = soybeanCombinedWASDE)
saveRDS(soybeanWASDE_clean, "preprocessed_data/soybeanWASDE_clean.Rds")
```
